---
title: "Customer Segmentation"
author: "Fadly"
date: "2024-08-07"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}
library(dplyr)
library(ggplot2)
library(gridExtra)  # For arranging multiple plots
library(fastDummies)
library(factoextra)
library(reshape2)
library(cluster)
```

```{r}
pelanggan <- read.csv("https://raw.githubusercontent.com/imomenujung/R-Projects/main/Multivariate-Analysis/Customer-Segmentation/Mall_Customers.csv", sep=",")
pelanggan
```

```{r}
str(pelanggan)
```

## Drop Customer ID
```{r}
df <- pelanggan %>% select(-CustomerID)
df
```

```{r}
# Plot for Age
p1 <- ggplot(df, aes(x=Age)) +
  geom_histogram(aes(y=..count..), binwidth=5, fill="skyblue", color="black", alpha=0.7) +
  geom_density(aes(y=..count..*5), color="blue", size=1) + 
  theme_minimal()

# Plot for Annual Income
p2 <- ggplot(df, aes(x=`Annual.Income..k..`)) +
  geom_histogram(aes(y=..count..), binwidth=10, fill="lightblue", color="black", alpha=0.7) +
  geom_density(aes(y=..count..*10), color="blue", size=1) + 
  theme_minimal()

# Plot for Spending Score
p3 <- ggplot(df, aes(x=`Spending.Score..1.100.`)) +
  geom_histogram(aes(y=..count..), binwidth=10, fill="slateblue", color="black", alpha=0.7) +
  geom_density(aes(y=..count..*10), color="blue", size=1) + 
  theme_minimal()

# Arrange plots in a grid
grid.arrange(p1, p2, p3, ncol=3)

```

```{r}
# Convert data to long format for plotting
df_long <- reshape2::melt(df)

# Create notched boxplot
ggplot(df_long, aes(x=value, y=variable, fill=variable)) +
  geom_boxplot(notch=TRUE, color="purple") +
  scale_fill_manual(values=c("skyblue", "lightblue", "blue")) +
  theme_minimal() +
  theme(legend.position="none")
```
Most of customers have age between 30 and 35. And the average age is 39.

The average annual income is arround 61 K$.
Most of customers have an spending score between 45 and 55. The average spending score is arround 50.
There is only one outliers for the variable "Annual Income".



```{r}
# Data for donut chart
gender_counts <- df %>% count(Gender)
gender_counts$percentage <- round(100 * gender_counts$n / sum(gender_counts$n), 2)

# Donut chart
donut_plot <- ggplot(gender_counts, aes(x=2, y=percentage, fill=Gender)) +
  geom_bar(stat="identity", width=1) +
  coord_polar(theta="y") +
  geom_text(aes(label=paste0(percentage, "%")), position=position_stack(vjust=0.5)) +
  xlim(0.5, 2.5) +
  theme_void() +
  theme(legend.position="none") +
  labs(title="Donut Chart") +
  theme(plot.title = element_text(hjust = 0.5, face="bold", color="blue")) +
  scale_fill_manual(values=c("lightblue", "blue"))

# Bar plot (Histogram)
histogram_plot <- ggplot(df, aes(x=Gender, fill=Gender)) +
  geom_bar(width=0.7, color="black", aes(y = ..count..)) +
  geom_text(stat='count', aes(label=..count..), vjust=-0.5, color="black") +
  theme_minimal() +
  labs(title="Histogram", y="Counts", x="Gender") +
  theme(plot.title = element_text(hjust = 0.5, face="bold", color="blue")) +
  scale_fill_manual(values=c("lightblue", "blue")) +
  theme(legend.position="none")

# Arrange the plots side by side
grid.arrange(donut_plot, histogram_plot, ncol=2)
```
As we can see 56% of customers are female, while 44% are mall.


```{r}
ggplot(df, aes(x = Age, y = Annual.Income..k.., color = Gender)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add regression line
  labs(title = "Age & Annual Income with line of best fit grouped by Gender",
       x = "Age", y = "Annual Income (k$)") +
  scale_color_manual(values = c("Male" = "#A63446", "Female" = "#17806D")) +  # Set colors
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  # Center title
```



```{r}
ggplot(df, aes(x = Age, y = Spending.Score..1.100., color = Gender)) +
  geom_point() +
  geom_smooth(method = "lm", se = TRUE) +  # Add regression line with confidence interval
  labs(title = "Age & Spending Score with line of best fit grouped by Gender",
       x = "Age", y = "Spending Score (1-100)") +
  scale_color_manual(values = c("Male" = "#A63446", "Female" = "#17806D")) +  # Set colors
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5)) +  # Center title
  theme(legend.position = "right")  # Adjust legend position
```

```{r}
ggplot(df, aes(x = Annual.Income..k.., y = Spending.Score..1.100., color = Gender)) +
  geom_point(alpha = 0.7) +  # Adjust alpha for point transparency
  geom_smooth(method = "lm", se = TRUE) +  # Add regression line with confidence interval
  labs(title = "Annual Income & Spending Score with line of best fit grouped by Gender",
       x = "Annual Income (k$)", y = "Spending Score (1-100)") +
  scale_color_manual(values = c("Male" = "#A63446", "Female" = "#17806D")) +  # Set colors
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5))  # Center title
```

```{r}
# Calculate correlation matrix

df_ng <- df %>% select(-Gender)
cor_matrix <- cor(df_ng)

# Melt the correlation matrix for ggplot
melted_cor <- melt(cor_matrix)

# Create the heatmap
ggplot(data = melted_cor, aes(Var2, Var1, fill = value)) +
  geom_tile(color = "white") +
  scale_fill_gradient2(low = "purple", high = "yellow", mid = "green", midpoint = 0, limit = c(-1, 1)) +
  geom_text(aes(Var2, Var1, label = round(value, 2)), color = "black", size = 3) +
  labs(title = "Correlation Heatmap", x = "", y = "") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

# Data Preprocessing
```{r}
# Select columns
col_names <- c("Annual.Income..k..", "Age", "Spending.Score..1.100.")
features <- df[, col_names]

# Standardize features (equivalent to StandardScaler)
scaler <- scale(features)  # Scales and centers features

# Create scaled data frame
scaled_features <- data.frame(scaler)  # Convert scaled matrix to data frame
colnames(scaled_features) <- col_names  # Set column names

scaled_features

```

```{r}
# Extract gender column
gender <- df$Gender

# Combine gender with scaled features
newdf <- bind_cols(scaled_features, Gender = gender)

# One-hot encoding
newdf <- newdf %>% 
  mutate(Gender = as.factor(Gender)) %>% 
  mutate(Gender = relevel(Gender, ref = "Male")) %>%  # Set reference level
  dummy_cols(select_columns = "Gender", remove_selected_columns = TRUE)

# Drop the first dummy variable

# Display the head of the DataFrame
head(newdf)
```


```{r}
newdf <- newdf %>% select(-Gender_Male)
newdf
```


# Building K-Means Clustering

```{r}
# Tentukan Cluster
jumlah_klaster <- c(1:9)  #Vektor yang berisikan jumlah klaster yang ingin dilihat nilai dari total within-cluster sum of squares
within_ss <- c()  #Vektor kosong yang akan diisi nilai total within-cluster sum of squares
for (i in jumlah_klaster) {
  within_ss <- c(within_ss, kmeans(x = newdf, centers = i, nstart = 25)$tot.withinss)
}

plot(x = jumlah_klaster, y = within_ss, type = "b", xlab = "Number of Cluster",
     ylab = "Total Within Sum of Squares", main = "Elbow Plot")
```

```{r}
# Tentukan Cluster
jumlah_klaster <- c(1:9)  #Vektor yang berisikan jumlah klaster yang ingin dilihat nilai dari total within-cluster sum of squares
within_ss <- c()  #Vektor kosong yang akan diisi nilai total within-cluster sum of squares
for (i in jumlah_klaster) {
  within_ss <- c(within_ss, pam(newdf, i)$tot.withinss)
}

plot(x = jumlah_klaster, y = within_ss, type = "b", xlab = "Number of Cluster",
     ylab = "Total Within Sum of Squares", main = "Elbow Plot")
```

```{r}
newdf <- filtered_df
# K-means
set.seed(123)
kmeans_clustering <- kmeans(x = newdf, centers = 4, nstart = 25)  #parameter nstart digunakan untuk memberitahu fungsi berapa kali inisiasi centroid awal (secara acak) yang akan dibentuk dan centers digunakan untuk memberitahu fungsi berapa jumlah klaster yang akan dibentuk.
kmeans_clustering

# Visualisasi Cluster
fviz_cluster(kmeans_clustering, data = newdf, 
             geom = "point",
             ellipse.type = "convex", 
             ggtheme = theme_bw()
)
```

```{r}
set.seed(42)  # Untuk reproduktibilitas
km_result <- eclust(newdf, "kmeans", k = 4 ,nstart = 42, graph = TRUE)  # Menggunakan K-Means++ dengan 3 cluster

# Visualisasi hasil clustering
fviz_cluster(km_result, geom = "point", ellipse.type = "norm",
             ggtheme = theme_minimal())
```

```{r}
set.seed(42)  # Untuk reproduktibilitas
kmedoids_result <- pam(newdf, k = 4)  # Menggunakan K-Medoids dengan 3 cluster

# Melihat hasil clustering
print(kmedoids_result)

# Melihat cluster yang dihasilkan
print(kmedoids_result$clustering)

fviz_cluster(kmedoids_result, geom = "point", ellipse.type = "norm",
             ggtheme = theme_minimal())
```

```{r}
set.seed(42)  # For reproducibility
km_result <- eclust(newdf, "kmeans", k = 4 ,nstart = 25, graph = TRUE) 

# Calculate the silhouette score
silhouette_score <- silhouette(km_result$cluster, dist(newdf))

# Calculate the average silhouette width (which is the silhouette score)
avg_silhouette_score <- mean(silhouette_score[, "sil_width"])

# Print the silhouette score
cat("Silhouette Score is:", avg_silhouette_score, "\n")
```

```{r}
boxplot(newdf)
```
```{r}
# Hapus Outlier
Q1 <- apply(newdf, 2, quantile, probs = 0.25)
Q3 <- apply(newdf, 2, quantile, probs = 0.75)
IQR <- Q3 - Q1

# Tentukan batas bawah dan atas untuk outlier
lower_bound <- Q1 - 1.5 * IQR
upper_bound <- Q3 + 1.5 * IQR

```

```{r}
filtered_df <- newdf
for(i in 1:ncol(newdf)) {
  filtered_df <- filtered_df[filtered_df[,i] >= lower_bound[i] & filtered_df[,i] <= upper_bound[i], ]
}

boxplot(filtered_df, main = "After Outlier Removal")
```


```{r}
# Melakukan PCA
pca_result <- prcomp(newdf, scale. = TRUE)

```

```{r}
explained_variance <- (pca_result$sdev)^2 / sum((pca_result$sdev)^2)
explained_variance
```

```{r}
# Menghitung varians kumulatif
cumulative_variance <- cumsum(explained_variance)
cumulative_variance
```

```{r}
# Membuat scree plot
plot(cumulative_variance, type = 'b', xlab = 'Number of Components', ylab = 'Cumulative Variance Explained', 
     main = 'Scree Plot')
abline(h = 0.9, col = 'red', lty = 2)  # Garis untuk threshold 90% varians
```

```{r}
# Menentukan jumlah komponen yang akan digunakan (misalnya 2 atau 3 komponen pertama)
pca_data <- data.frame(pca_result$x[, 1:3])  # Menggunakan 2 komponen pertama
```

```{r}
# Melakukan clustering pada data PCA
set.seed(42)  # Untuk reproduktibilitas
kmeans_model <- kmeans(pca_data, centers = 4, nstart = 25)  # Misalnya 3 cluster
```

```{r}
silhouette_score <- silhouette(kmeans_model$cluster, dist(pca_data))
avg_silhouette_score <- mean(silhouette_score[, "sil_width"])
print(avg_silhouette_score)
```

```{r}
set.seed(42)  # Untuk reproduktibilitas
kmedoids_result <- pam(pca_data, k = 4)  # Menggunakan K-Medoids dengan 3 cluster

# Melihat hasil clustering
print(kmedoids_result)

# Melihat cluster yang dihasilkan
print(kmedoids_result$clustering)

fviz_cluster(kmedoids_result, geom = "point", ellipse.type = "norm",
             ggtheme = theme_minimal())
```
```{r}
silhouette_score <- silhouette(kmedoids_result$cluster, dist(pca_data))
avg_silhouette_score <- mean(silhouette_score[, "sil_width"])
print(avg_silhouette_score)
```
